### 2013-07-25 main/s_level_0.tex
fix makefile; add section "Level 0" file; change wording on strictness

\section{Equivalence in the absence of $\parr$}