### 2014-01-13 main/mllequiv.tex
final lemma completing the puzzle

\documentclass{robinminion}

\usepackage{amsthm}
\usepackage{enumitem}
\usepackage[round]{natbib}
\usepackage{mllequiv,willemtools}

\graphicspath{{../notes/}}

\makeTheoremDefs

\author{Willem Heijltjes and Robin Houston}
\title{The proof equivalence problem for multiplicative linear logic is \textsc{pspace}-complete v0.3}

\begin{document}

\maketitle

\begin{abstract}
MLL proof equivalence is the problem of deciding whether two proofs in multiplicative linear logic are related by a series of inference permutations.
%
Previous work has shown the problem to be equivalent to a rewiring problem on proof nets, which are not canonical for full MLL due to the presence of the two units.
%
Drawing from recent work on reconfiguration problems, in this paper it is shown that MLL proof equivalence is PSPACE-complete, using a reduction from Nondeterministic Constraint Logic.
\end{abstract}

%\include{s_intro}
\include{s_background}
\include{s_level_0}
\include{s_level_1}
\include{s_level_2} 
%\include{s_conclusion}

\bibliographystyle{plainnat}
\bibliography{refs}

\end{document}