### 2014-01-20 main/scratch.tex
\aftersubsection

\documentclass[conference]{IEEEtran}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{mllequiv,willemtools}

\makeTheoremDefs

\begin{document}

\[

\]

\end{document}