### 2015-06-24 Reviews LMCS/Response.txt
many updates

We thank the referees for their careful reading of our paper, and for the detailed and helpful feedback they have provided. Here, we will discuss each review in turn, as well as the changes we have made in response. We omit comments whose

Review 1

p.1, l.-1: "Canonical proof nets may also be defined for intuitionistic MLL with units": are you referring to (some of) the same papers you cite in the next paragraphs (i.e. Blute et al., Trimble, Lamarche, etc.)?  In any case, you should add a reference here.

*** We have tried and failed to find a decent reference for the kind of proof nets we have in mind here, although private discussions have convinced us that the issue is quite widely understood. The principle is simple: re-wiring jumps towards the unique conclusion yields a canonical form. But Lamarche's "essential nets", extended with units by Murawski and Ong (both of which we have added as a reference), attach jumps to the leaves of the tree instead.

p.10, notations: why do you use a "prime" in your notation for sequents/formulas?  I mean, why n' (in a circle) instead of just n?  Since n',i',j', etc. are common notations for variables (as n,i,j...), the notation is quite misleading at first.

*** The prime is used as a notation for 'successor', to indicate that a node labelled n' has n + 1 bottoms or ones, while n' and m' connect to (n+m)'.

p.25, Definition 37: I find this definition a bit disturbing, it seems ad hoc at first and leaves one wondering whether something is wrong with your encoding. [...]

*** We agree that the purpose of Definition 37 is not well-explained. We have opted for an orthogonal solution: we now explicitly define the translation of an instance of NCL-reconfiguration (which consists of a constraint graph with two configurations).

p.30, Theorem 46: just a minor quibble: you could make explicit the reduction from NCG-reconfiguration to the proof net equivalence problem, in order for the reader to better understand the complexity of such a reduction, which must be at least polytime computable. [...]

 ??? We have also made the complexity of this translation more explicit (in Theorem 46).

Review 2

I have the impression that the authors do not realize the importance of their own result.
From the point of view of the theory of proof nets, the problem concerns how to deal with lack of connectedness. It is formulated in terms of the multiplicative units but this is just for technical convenience. The reason why the problem is relevant is the fact that the bottom unit is a variant of the weakening rule, one of the cornerstones of linear logic. The multiplicative units can indeed be easily encoded by the exponentials (fix a type A and let 1 be encoded as an axiom on A, whose conclusions are connected by a par and then wrapped in a promotion, i.e. !(A^\perp \par A), and let bottom be the weakening on the dual formula). Therefore PSPACE­-completeness of proof equivalence for MLL implies PSPACE­hardness of proof equivalence for MELL. This is the real issue, as the exponentials are the expressive core of linear logic. Essentially, the paper proves that proof nets for MELL are intractable, unless ad-­hoc notions as the mix rules are added, or intractability is encapsulated in the translation from proofs to nets. Such a consequence is not explained in the paper, while there is no doubt that it has to be mentioned as it is the main reason of interest for the result.

Another issue concerns the paragraph of the introduction that mentions the fragments of linear logic for which canonical proof nets are known. I think that the polarized case should be added. Moreover it is said that Intuitionistic MLL with units has a canonical presentation but
no citations are given. Of course, *a citation has to be given*, but that is not the point. Given the results in the paper, Intuitionistic (ME)LL is the only hope for a tractable and expressive theory of proof nets. The intuitionistic case should then be discussed precisely, explaining the differences with the classical case. The canonical presentation of the intuitionistic case that the authors have in mind probably relies on jumps, and thus rewiring and proof equivalence make sense there too. Why is that case tractable? And does tractability rely on the fact that cut­-elimination in the linear case is trivial? Would Intuitionistic MELL be tractable as well?

I also find the third paragraph of page 2 (“Since there can be no canonical representation...”) unsatisfying. It points to alternative ideas in the literature to define the identity of proofs. They have to be discussed, at least briefly. At present the alternative approaches are just cited, without giving any hint of why they would help to treat the units.
I am also surprised by the total absence of a discussion of the case of MLL with the MIX rules.
A better treatment of the context of the problem cannot be avoided, because it is the reason why the result is interesting. Moreover, it is also the reason that justifies to not take too seriously the many technicalities of the

Review 3

- p.21: the bijection associates f(1) to j and c(3) to f

*** In fact, the association f(1)-f and c(3)-c in the paper is correct: each edge from parent to child is associated with the child.