### 2015-07-12 Reviews LMCS/Response.txt
more edits; final author response

We thank the referees for their careful reading of our paper, and for the detailed and helpful feedback they have provided. Here, we will discuss each review in turn, and highlight the changes we have made in response. For the suggested minor changes, we have implemented most, and will comment only on those we chose not to.

Review 1

p.1, l.-1: "Canonical proof nets may also be defined for intuitionistic MLL with units": are you referring to (some of) the same papers you cite in the next paragraphs (i.e. Blute et al., Trimble, Lamarche, etc.)?  In any case, you should add a reference here.

*** We have tried and failed to find a decent reference for the kind of proof nets we have in mind here, although private discussions have convinced us that they are quite widely understood. The principle is simple: re-wiring jumps towards the unique conclusion yields a canonical form. But Lamarche's "essential nets", extended with units by Murawski and Ong (both of which we have added as a reference), attach jumps to the leaves of the tree instead.

p.10, notations: why do you use a "prime" in your notation for sequents/formulas?  I mean, why n' (in a circle) instead of just n?  Since n',i',j', etc. are common notations for variables (as n,i,j...), the notation is quite misleading at first.

*** We have removed the primes.

p.25, Definition 37: I find this definition a bit disturbing, it seems ad hoc at first and leaves one wondering whether something is wrong with your encoding. [...]

*** We agree that Definition 37 is not well-explained. We have opted for the following solution: we now explicitly define the translation of an instance of NCL-reconfiguration (which consists of a constraint graph with two configurations).

p.30, Theorem 46: just a minor quibble: you could make explicit the reduction from NCG-reconfiguration to the proof net equivalence problem, in order for the reader to better understand the complexity of such a reduction, which must be at least polytime computable. [...]

*** We have also made this translation more explicit (in the final theorem as well as earlier in the paper).

Review 2

I have the impression that the authors do not realize the importance of their own result. [...]

*** While we were aware that MELL (without units) may encode MLL, we indeed did not realise that this consequence of our result (that MELL has no canonical normal forms) would be unexpected, or considered more interesting than the consequence for MLL. And we remain slightly cautious: surely proof equivalence for MELL is completely dominated by the complexity of cut-elimination? We have added a paragraph to this effect to the introduction.

Another issue concerns the paragraph of the introduction that mentions the fragments of linear logic for which canonical proof nets are known. I think that the polarized case should be added.

*** We have now mentioned more explicitly that polarized proof nets are canonical, further down from that paragraph.

Moreover it is said that Intuitionistic MLL with units has a canonical presentation but no citations are given. [...]

*** We have expanded the explanation of IMLL proof nets that we have in mind. (See also above in the response to Review 1.) We view intuitionistic LL as one of several alternatives to classical LL (as remarked on in the next comment).

I also find the third paragraph of page 2 (“Since there can be no canonical representation...”) unsatisfying. It points to alternative ideas in the literature to define the identity of proofs. They have to be discussed, at least briefly. [...]

*** We have expanded the discussion of these alternatives, giving brief hints as to how they avoid the unit problem.

I am also surprised by the total absence of a discussion of the case of MLL with the MIX rules.

*** We opted not to discuss MIX - depending on which rules are added, either the units are equated, which is semantically equivalent to omitting them, or they are not, in which case the logic is conservative over MLL and the unit problem remains.

1) Numerical constraints [...]

*** We have added a further illustration showing where the multiplication by 3x the number of edges is needed.

2) Parity [...]

*** We have added some introductory paragraphs to the subsection on parity, and adjusted Definition 37 for clarity (see also the response to Review 1 above).

3) Soundness/readback[...]

*** We have put the read-back or decoding function in a separate proposition, with justification. The only concern for decoding is the following lemma: if two linkings are equivalent, there is a reconfiguration path between their decodings.

First, the authors choose to restrict jumps to target 1-­formulae and to rewire jumps according to a big-­step strategy, building on the work of Hughes. [...]

*** We now have more clearly motivated this choice: it is mostly cosmetic, as it simplifies the diagrammatic notation for proof nets - though it does impact the formulation of some lemmas in the section "Rewiring proof nets".

Second, in Section 4 it is said that the reconfiguration problem remains PSPACE-­complete under a number of restrictions, e.g. fixing constraint at the value 2, but the authors choose the encode the general case, without restrictions. [...]

*** We don't think these restrictions would actually simplify the encoding - in fact, it would be exactly the same. It would still be necessary, for instance, to correct for parity, and to multiply weights by 3 times the number of edges.

Review 3

- p.21: the bijection associates f(1) to j and c(3) to f

*** In fact, the association f(1)-f and c(3)-c in the paper is correct: each edge from parent to child is associated with the child. We have amended the formulation of this bijection for clarity.